---
title: "wine"
author: "Elisabetta Rocchetti"
date: "15/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(caret)
library(ggplot2)
```

# Data

The datasets **white_wines.csv** and **red_wines.csv** store a bunch of data explaining physicochemical properties of many wines which come from the north-west region, named Minho, of Portugal.

Each row refers to a wine and each column contain one of the physicochemical properties with respect to wines. Here we list there properties:

- fixed acidity (FA: g(tartaric acid)/dm^3)
- volatile acidity (VA: g(acetic acid)/dm^3)
- citric acid (CA: g/dm^3)
- residual sugar (RS: g/dm^3)
- chlorides (CH: g(sodium chloride)/dm^3)
- free sulfur dioxide (FSD: mg/dm^3)
- total sulfur dioxide (TSD: mg/dm^3)
- density (DE: g/dm^3)
- pH
- sulphates (SU: g(potassium sulphate)/dm^3)
- alcohol (AL: %vol)
- quality (from 0 to 10)

## Load the datasets

```{r, message=FALSE, warning = FALSE}
winequality.white <- read.csv("winequality-white.csv", sep=";")
winequality.red <- read.csv("winequality-red.csv", sep=";")
print("White wines")
str(winequality.white)
head(winequality.white)
print("Red wines")
str(winequality.red)
head(winequality.red)
print("White wines: ")
nrow(winequality.white)
print("Red wines: ")
nrow(winequality.red)
```
Since we want a unified dataset containing both red ans white wines, in the following we merge the two datasets adding a column representing the type of all wines: this variable is a dummy which gives 0 to red wines and 1 to white wines. We also balance the final dataset to have equal number of red and white wines.

## Adding 'type' variable, balancing and merging datasets

```{r, message=FALSE, warning = FALSE}
#adding type column
winequality.red$type <- rep(0,nrow(winequality.red))
winequality.white$type <- rep(1,nrow(winequality.white))

#balancing: reduce number of white wines
winequality.white <- sample_n(winequality.white, nrow(winequality.red))

#combining the two datasets
wines <- rbind(winequality.red,winequality.white)

print("All wines")
str(wines)
head(wines)
```
Now we inspect the variables we have and we transform them if necessary.

## Data preparation

First thing to fix: type variable represent a class, not a numeric discrete variable. Let's transform it into a factor.

```{r, message=FALSE, warning = FALSE}
wines$type <- as.factor(wines$type)
wines$quality <- as.factor(wines$quality)
```

In order to have a clearer interpretation of results, we transform each variable to cg/dm^3.

```{r, message=FALSE, warning = FALSE}
#from grams to centigrams
wines$fixed.acidity <- wines$fixed.acidity*10
wines$volatile.acidity <- wines$volatile.acidity*10
wines$citric.acid <- wines$citric.acid*10
wines$residual.sugar <- wines$residual.sugar*10
wines$chlorides <- wines$chlorides*10
wines$density <- wines$density*10
wines$sulphates <- wines$sulphates*10

#from milligrams to centigrams
wines$free.sulfur.dioxide <- wines$free.sulfur.dioxide/10
wines$total.sulfur.dioxide <- wines$total.sulfur.dioxide/10
```

Now that we have prepared the data, let's describe it.

```{r, message=FALSE, warning = FALSE}
summary(wines)
boxplot(wines$fixed.acidity ~ wines$type)
boxplot(wines$volatile.acidity ~ wines$type)
boxplot(wines$citric.acid ~ wines$type)
boxplot(wines$residual.sugar ~ wines$type)
boxplot(wines$chlorides ~ wines$type)
boxplot(wines$free.sulfur.dioxide ~ wines$type)
boxplot(wines$total.sulfur.dioxide ~ wines$type)
boxplot(wines$density ~ wines$type)
boxplot(wines$pH ~ wines$type)
boxplot(wines$sulphates ~ wines$type)
boxplot(wines$alcohol ~ wines$type)

ggplot(wines, aes(x = quality)) +
  geom_bar(aes(fill = type)) +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)

ggplot(wines, aes(x = type)) +
  geom_bar(aes(fill = type)) +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)
```


# Divide the dataset in training and test
```{r}
set.seed(42)
split_train_test <- createDataPartition(wines$type,p=0.7,list=FALSE)
dtrain<- wines[split_train_test,]
dtest<-  wines[-split_train_test,]
```

# Logistic regression

```{r, message=FALSE, warning=FALSE}
lr_fit <- glm(type ~., data = dtrain,
          family=binomial(link='logit'))
summary(lr_fit)
```
Computing different thresholds.
```{r, warning=FALSE}
exp(cbind(OR = coef(lr_fit), confint(lr_fit)))
```
Removing quality from our model.
```{r}
lr_fit_noquality <- glm(type ~fixed.acidity+volatile.acidity+citric.acid+residual.sugar+chlorides+free.sulfur.dioxide+total.sulfur.dioxide+density+pH+sulphates+alcohol, data = dtrain,
          family=binomial(link='logit'))
summary(lr_fit_noquality)
```
```{r}
exp(cbind(OR = coef(lr_fit_noquality), confint(lr_fit_noquality)))
```


**Confusion Matrix at various thresholds - test set**
```{r}
lr_prob1 <- predict(lr_fit, dtest, type="response")
lr_preds_test <- c(0,0,0,0,0,0,0,0,0,0,0)
i<-1
for (thresh in seq(0.25,0.75,0.05)){
  lr_pred <- ifelse(lr_prob1 > thresh,1,0)
  cm <- confusionMatrix(
          as.factor(lr_pred),
          as.factor(dtest$type),
          positive = "1"
          )
  lr_preds_test[i] <- cm$overall[[1]]
  i<-i+1
}
names(lr_preds_test) <- seq(0.25,0.75,0.05)
#lr_pred1 <- ifelse(lr_prob1 > 0.5,1,0)
#table(Predicted = lr_pred1, Actual = dtest$type)
```

**Confusion Matrix at various thresholds - training set**
```{r}
lr_prob2 <- predict(lr_fit, dtrain, type="response")
lr_preds_train <- c(0,0,0,0,0,0,0,0,0,0,0)
i<-1
for (thresh in seq(0.25,0.75,0.05)){
  lr_pred <- ifelse(lr_prob2 > thresh,1,0)
  cm <- confusionMatrix(
          as.factor(lr_pred),
          as.factor(dtrain$type),
          positive = "1"
          )
  lr_preds_train[i] <- cm$overall[[1]]
  i<-i+1
}
names(lr_preds_train) <- seq(0.25,0.75,0.05)
#lr_pred1 <- ifelse(lr_prob1 > 0.5,1,0)
#table(Predicted = lr_pred1, Actual = dtest$type)
```

**Accuracy**
```{r}
lr_prob2 <- predict(lr_fit, dtrain, type="response")
lr_pred2 <- ifelse(lr_prob2 > 0.5,1,0)
lr_tab1 <- table(Predicted = lr_pred2, Actual = dtrain$type)
lr_tab2 <- table(Predicted = lr_pred1, Actual = dtest$type)

# Train
confusionMatrix(
  as.factor(lr_pred2),
  as.factor(dtrain$type),
  positive = "1"
)

# Test
confusionMatrix(
  as.factor(lr_pred1),
  as.factor(dtest$type),
  positive = "1"
)
```

# ROC Curve

```{r, message=FALSE, warning=FALSE}
library(pROC)
lr_prob2 <- predict(lr_fit, dtest, type="response")
test_roc = roc(dtest$type ~ lr_prob2, plot = TRUE, print.auc = TRUE)
as.numeric(test_roc$auc)
```
# Relation with residual.sugar 

```{r}
ggplot(wines, aes(x = residual.sugar, fill = type)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Residual sugar",
       title = "Wine type by residual sugar")+
  xlim(0,25)

```

# Relation with chlorides

```{r}
ggplot(wines, aes(x = chlorides, fill = type)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Residual sugar",
       title = "Wine type by chlorides")

```
# Relation with total.sulfur.dioxide

```{r}
ggplot(wines, aes(x = total.sulfur.dioxide, fill = type)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Residual sugar",
       title = "Wine type by total.sulfur.dioxide")

```
# Relation with alcohol 

```{r}
ggplot(wines, aes(x = alcohol , fill = type)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Residual sugar",
       title = "Wine type by alcohol")

```